---
title: Basic crossed-hands analyses
author: "Kaian Unwalla"
date: "November, 2021"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

This code first summarizes the raw data, provides a couple different ways to visualize the summarized data, and finally conducts some frequentist statistics on the crossed-hands data. 

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.align = 'center', fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE, include = TRUE)
```

```{r, libraries}
rm(list=ls())
library(tidyverse)
library(rstatix)

source("Model functions.R")

time<-format(Sys.time(),"%b%d_%H%M%S")
path = "Data/"

alldata1 <- read.table(paste(path,"Test Data.txt", sep = ""), header = T)
```

```{r get_data}
#This function serves as a pivot table and calculates the mean of the the proportion right-first responses, based on the condition, hand posture, and SOA.
overall_lines <- Pivot_overall(alldata1) 

#This uses the pivot function which does the same thing as pivot_overall, but calculates it separately for each participant. 
indiv_lines <- Pivot(alldata1) %>% 
  select(-c(numTrials,numRF))

#Another analysis often conducted on crossed-hands data is the proportion correct difference (PCD) score. This score takes the difference between participants' crossed and uncrossed performance at each SOA and sums them together. This results in a single number that essentially indexes the accuracy cost for crossing the hands. A larger score indicates worse performance (larger crossed-hands deficit)
overall_PCD <- PCD_error(alldata1) 

#This uses the PCD function which does the same thing as PCD_error, but calculates it separately for each participant.
indiv_PCD <- PCD(alldata1) 
```

## Graphs
This code creates an overall graph of participant performance. 
```{r overall_graph}
#plots overall data separated by hand posture and condition
overall_lines %>% 
  ggplot(aes(x = SOA, y = Data, group = interaction(Condition,Hands), linetype = Hands, fill = Condition))+
  geom_errorbar(aes(ymin=Data-SE,ymax=Data+SE),width=.01, linetype = 'solid')+
  geom_line(size=0.5)+geom_point(size=3, shape=21)+
  scale_linetype_manual(values = c('dashed','solid'))+
  scale_fill_manual(values = c("grey25", "grey"))+
  scale_x_continuous(breaks=seq(-0.4,0.4,0.4))+
  scale_y_continuous(breaks=seq(0,1,0.5))+
  ylab("Proportion of 'Right-First' Responses")+xlab("SOA (s)")
```
- You can see that performance is almost equivalent across the two conditions when the hands are uncrossed. When the hands are crossed, participants are performing slightly better in condition A, and close to chance in condition B. 

This code here plots the individual participants, so you can see what is happening for each participant separately. 
```{r, data_graphs}
#plots line graph, with performance separated by hand posture and condition for each participant
indiv_lines %>%
  ggplot(aes(x = SOA, y = Data, group = interaction(Condition,Hands), linetype = Hands, color = Condition))+
  geom_line()+
  scale_linetype_manual(values = c('dashed','solid'))+
  scale_color_manual(values = c("grey25", "grey"))+
  scale_x_continuous(breaks=seq(-0.4,0.4,0.4))+
  scale_y_continuous(breaks=seq(0,1,0.5))+
  ylab('Proportion of Right First Responses')+xlab("SOA (ms)")+
  facet_wrap(~Participant)
```

This plots the overall PCD score for each condition. 
```{r overall_PCD, fig.width=4, fig.height=5, include=T}
overall_PCD %>% 
  ggplot(aes(x=Condition,y=PCD,fill=Condition)) + 
  geom_bar(stat="identity",position=position_dodge(), width = 0.5) + 
  scale_fill_manual(values = c('grey25','grey'))+
  geom_errorbar(aes(ymin=PCD-SE,ymax=PCD+SE),width=.05)+
  ylab('PCD Score')  
```
- You can see that condition A has a smaller PCD score than condition B. This supports what we see in the individual line graphs where there is a smaller separation between the uncrossed and crossed lines for condition A compared to condition B. 


This code here plots the individual participants' PCD scores, so you can see what is happening for each participant separately. 
```{r individual_PCD, include=T}
indiv_PCD %>% 
  ggplot(aes(x=Condition ,y=PCD, fill = Condition)) + 
  scale_fill_manual(values = c('grey25','grey'))+
  geom_bar(stat='identity', position = position_dodge(),width = 0.8)+
  facet_wrap(~Participant)
```


##Stats
```{r, include=TRUE}
#convert participant and SOA to factors for the ANOVA 
indiv_lines$Participant <- factor(indiv_lines$Participant)
indiv_lines$SOA <- factor(indiv_lines$SOA)

#calculate anova using anova_test from rstatix package
indiv_aov <- anova_test(data = indiv_lines, dv = Data, wid = Participant, within = c(Condition,Hands,SOA))
get_anova_table(indiv_aov)
```
- Based on the results of the anova we can see a significant effect of SOA, suggesting that performance changes as the time difference between the two vibrations changes.
- There is also an interaction between condition and SOA, showing that the change in SOA is different for condition A and B
- There is also an interaction between hand posture and SOA, showing that the change in SOA is different when the hands are crossed and uncrossed. This indicates that participants are performing worse when their hands are crossed. 
- The interaction of interest is the significant interaction between condition, hand, and SOA. This suggests that the worse performance seen when the hands are crossed is different for condition A and condition B. 

A simpler statistic is to simply do a t-test on the PCD scores to see whether there is a change in the magnitude based on the condition. 
```{r, order_anova, include=T}
indiv_PCD$Participant <- factor(indiv_PCD$Participant)
t.test(PCD~Condition, data = indiv_PCD)
```
- Based on the data it is evident that condition B has a smaller PCD score, and therefore a smaller crossed-hands deficit. This suggests that the manipulation used for condition B helped reduce the conflict between internal and external information and improved tactile localization. 

